Predicting prolonged apnea during nurse-administered procedural sedation using machine learning

### 2021-03-26 manuscript.Rmd
load data from figshare

---
title: "Predicting prolonged apnea during nurse-administered procedural sedation using machine learning"
bibliography: "Extras/references.bib"
csl: "Extras/anaesthesia.csl"
author:
  - Aaron Conway PhD:
      email: aaron.conway@utoronto.ca
      phone: +1 416 946 7112
      address: 585 University Ave | Toronto, ON  M5G 2N2
      institute: [PMCC, LBFON, QUT]
      correspondence: true
  - Carla Jungquist PhD:
      institute: [UoB]
      correspondence: false
  - Kristina Chang MScN:
      institute: [PMCC]
      correspondence: false
  - Navpreet Kamboj BScN:
      institute: [LBFON]
      correspondence: false
  - Joanna Sutherland MBBS:
      institute: [NSW, Coffs]
      correspondence: false
  - Sebastian Mafeld:
      institute: [IR]
      correspondence: false
  - Matteo Parotto MD, PhD:
      institute: [TGH, UoT]
      correspondence: false
institute:
  - PMCC: Peter Munk Cardiac Centre, University Health Network, Toronto, Canada
  - LBFON: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto, Toronto, Canada
  - QUT: School of Nursing, Queensland University of Technology (QUT), Brisbane, Australia
  - UoB: The University of Buffalo, Buffalo, New York, USA
  - NSW: Rural Clinical School, University of New South Wales, Coffs Harbour, NSW, Australia
  - Coffs: Department of Anaesthesia, Coffs Harbour Health Campus, Coffs Harbour, NSW, Australia
  - IR: Interventional Radiology, JDMI, Toronto General Hospital
  - TGH: Department of Anesthesia and Pain Management, Toronto General Hospital, UHN, Toronto, Canada
  - UoT: Department of Anesthesiology and Pain Medicine and Interdepartmental Division of Critical Care Medicine, University of Toronto, Toronto, Canada
output:
  officedown::rdocx_document:
    reference_docx: Extras/manuscript_template.docx
    pandoc_args:
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
---

Declarations of interest: none

Funding: A.C. received support from a National Health and Medical Research Council Early Career Fellowship (APP1091657) and the Bertha Rosenstadt Faculty Small Research Grant from the Lawrence S. Bloomberg Faculty of Nursing at the University of Toronto. This study received funding from the Wesley Medical Research (Project number: 2015-31).

Role of the funding source: The funders had no role in study design; in the collection, analysis and interpretation of data; in the writing of the report; and in the decision to submit the article for publication.

@@

@@

@@

@@

## Conclusion

@@

Deciphering which capnography waveform abnormalities deserve intervention (and therefore alarms to signal the event to clinicians) from those that do not is an essential step towards successfully implementating this technology into practice. For example, triggering alarms after short periods of apnea leads to frequent interruptions and potentially increases the risk of alarm fatigue. Conversely, only intervening once an apneic period reaches a longer threshold negates capnography's potential benefits on patient safety by improving ventilation. In practice, two alternative strategies for capnography alarm management are typically used. The 'aggressive' strategy is for alarms to be triggered after short periods of apnea (e.g. 15 seconds). The 'conservative' approach triggers the alarm only once the patient has been apneic for a prolonged period (e.g. 30 seconds).

@@

Analyses were performed using R[@rcore2021]. Data as well as details about how to access the code and a reproducible computing environment required to verify the results is available from this webpage: <https://github.com/awconway/capnopred>.

>>>>

---
title: "Predicting prolonged apnea during nurse-administered procedural sedation using machine learning"
bibliography: "Extras/references.bib"
csl: "Extras/anaesthesia.csl"
# author:
#   - Aaron Conway PhD:
#       email: aaron.conway@utoronto.ca
#       phone: +1 416 946 7112
#       address: 585 University Ave | Toronto, ON  M5G 2N2
#       institute: [PMCC, LBFON, QUT]
#       correspondence: true
#   - Carla Jungquist PhD:
#       institute: [UoB]
#       correspondence: false
#   - Kristina Chang MScN:
#       institute: [PMCC]
#       correspondence: false
#   - Navpreet Kamboj BScN:
#       institute: [LBFON]
#       correspondence: false
#   - Joanna Sutherland MBBS:
#       institute: [NSW, Coffs]
#       correspondence: false
#   - Sebastian Mafeld:
#       institute: [IR]
#       correspondence: false
#   - Matteo Parotto MD, PhD:
#       institute: [TGH, UoT]
#       correspondence: false
# institute:
#   - PMCC: Peter Munk Cardiac Centre, University Health Network, Toronto, Canada
#   - LBFON: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto, Toronto, Canada
#   - QUT: School of Nursing, Queensland University of Technology (QUT), Brisbane, Australia
#   - UoB: The University of Buffalo, Buffalo, New York, USA
#   - NSW: Rural Clinical School, University of New South Wales, Coffs Harbour, NSW, Australia
#   - Coffs: Department of Anaesthesia, Coffs Harbour Health Campus, Coffs Harbour, NSW, Australia
#   - IR: Interventional Radiology, JDMI, Toronto General Hospital
#   - TGH: Department of Anesthesia and Pain Management, Toronto General Hospital, UHN, Toronto, Canada
#   - UoT: Department of Anesthesiology and Pain Medicine and Interdepartmental Division of Critical Care Medicine, University of Toronto, Toronto, Canada
output:
  officedown::rdocx_document:
    reference_docx: Extras/manuscript_template.docx
    # pandoc_args:
    #   - --lua-filter=Extras/scholarly-metadata.lua
    #   - --lua-filter=Extras/author-info-blocks.lua
---

@@

### Introduction

@@

### Methods

@@

### Results

@@

@@

Deciphering which capnography waveform abnormalities deserve intervention (and therefore alarms to signal the event to clinicians) from those that do not is an essential step towards successfully implementating this technology into practice. For example, triggering alarms after short periods of apnea leads to frequent interruptions and potentially increases the risk of alarm fatigue. Conversely, only intervening once an apneic period reaches a longer threshold negates capnography's potential benefits on patient safety by improving ventilation. In practice, two alternative strategies for capnography alarm management are typically used. The 'aggressive' strategy is for alarms to be triggered after short periods of apnea (e.g. 15 seconds). The 'conservative' approach triggers the alarm only once the patient has been apneic for a prolonged period (e.g. 30 seconds). Preferences for the aggressive or conservative alarm threshold will be influenced by many factors, including the rate of oxygen supplmentation. The duration of time between the onset of apnea to hypoxemia increases with higher flows of oxygen.[@wong2019high]

@@

Analyses were performed using R[@rcore2021]. Data as well as details about how to access the code and a reproducible computing environment required to verify the results is available from this webpage: <https://github.com/awconway/capnopred>. The raw data is archived here: https://figshare.com/account/projects/101195/articles/14320874.
